Classifying galaxy spectra at $0.5<z<1$: an unsupervised approach

### 2016-06-23 sections0.05/results0.05.tex
Update results0.05.tex
small edits. No new questions.

    In order to find a sufficient size for the trained networks, we created maps with sizes ranging from $1\times2$ to $50\times50$.
    Varying the grid size of the maps helps us to monitor whether tighter grouping of galaxies is due to their similar properties or a lack of map space to separate them.
    Based on the size of the data and SOM results, we found the optimum grid size to be $1\times22$ and $12\times12$ in 1D and 2D maps, respectively. 
    For each grid we created different SOMs with different learning factors, neighbourhood distances, and number of iterations to find the most suitable values for our dataset, which are listed in Section~\ref{sec: create_som}. %%%Sr220616: changed the optimum to the most suitable values ... . moved the initial values to section 3.2

@@

    \subsection{One-dimensional self-organizing maps}
    \label{sec: 1D}
        \subsubsection{TRAINING THE NETWORKS}
        \label{sec: 1Dt}
            To start our clustering, we assumed that galaxies can be divided into only two general types; quiescent and starburst.
            This corresponds to a network with only two neurons.
            We increased the size of the map gradually until the 12 input samples divide into the 12 neurons.

            Figs.~\ref{fig: 1by2T} --\ref{fig: 1by22T}~show the results of the training $1\times2$, $1\times3$, and $1\times22$ networks, respectively (The rest of the networks are shown in Appendix~\ref{app: high_Z_1d_soms}.
            As in to Fig.~\ref{fig: sample}, the upper part of the figures shows the neurons and their relative distances, $D_j$.
            As mentioned in Sec.~\ref{sec: method}, an increase in the darkness of colours between neurons represents an increase in relative distance between the neurons.
            The lower panels of the figures show the number of \citetalias{Kinney96} templates that are placed in each neuron. 
            \begin{figure}
                \begin{subfigure}[b]{0.5\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{images0.01/1d/dist_1_by_2.png}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.5\textwidth}
                     \includegraphics[width=\textwidth]{images0.01/1d/hit_t_1_by_2.png}
                \end{subfigure}
                \caption{Results of training network in $1\times2$~grid. As in Fig.~\ref{fig: sample}, the upper panel is a distance map and the lower panel is a hit map. In this network, 5 of the templates from \citetalias{Kinney96} are categorized as quiescent galaxies and the rest are starbursts. Because of their strong emission lines, Sc galaxies are moved towards the starburst ones.}
                 \label{fig: 1by2T}
            \end{figure}

            In the upper map in Fig.~\ref{fig: 1by2T}, the dark colour between two neurons indicates that the relative distance between these two groups are relatively high, and these two groups are distinguishable groups.
            In the lower part of Fig.~\ref{fig: 1by2T}, we see that the templates are divided into two groups of 5 and 7.
            Although we know from \citetalias{Kinney96} and \citetalias{Hossein12} that 6 of the templates are quiescent and the other 6 are starbursts, the SOM results show 5 of the galaxies in one group and the other 7 in the second group.
            In this method, the Sc template has been categorized as starburst due to the relatively higher disk stellar population relative to that of the bulge. 
            According to \citetalias{Kinney96}, Sc galaxies are considered to be late Hubble type galaxies which have a flattter SED compare to other quiescent galaxies.

            Fig.~\ref{fig: 1by3T} shows the results of the training in a 1$\times$3 network.
            In these plots we force the galaxies to be categorize in a maximum 3 groups. 
            If the templates in Fig.~\ref{fig: 1by2T} truly belonged in two groups, they would be grouped into two groups in this network, even when we try to cluster them into three. 
            However, in the lower panel of Fig.~\ref{fig: 1by3T}, we can see that the middle node contains two templates.
            These two templates, which are separated from the group of starburst templates in the lower part of Fig.~\ref{fig: 1by2T},  are the SB5 and SB6 types.
            In the upper plot of Fig.~\ref{fig: 1by3T}, the colour between two right neurons is black and the colour between two left neurons is white. 
            The black colour indicates that the left neuron is completely different from the other two groups,
            while the white colour shows that the two right neurons are very similar to one another.

@@

            We increased the size of the maps gradually until the galaxies are divided into twelve groups (Figs.~\ref{fig: 1by4T} to ~\ref{fig: 1by20T} and ~\ref{fig: 1by22T}).
            Since there are more nodes in higher grid SOMs, the  algorithm 
            pays more attention to small differences between groups.
            If the templates from \citetalias{Kinney96} had completely distinct SEDs, then a $1\times12$-sized SOM would be expected to show 12 different groups each containing a single template.
            However, in Fig.~\ref{fig: 1by12T}, we see that three of the neurons contain two templates and three of them are empty.
            It is evident that there was no template in the \citetalias{Kinney96} sample that can fill those empty neurons.
            In Fig.~\ref{fig: 1by12T}, from left to right, templates with types B and E, SB3 and SB4, and SB1 and SB2 are the ones grouped together. 
            The SB3 and SB4 grouping breaks when we increase the size of the network to $1\times15$~(Fig.~\ref{fig: 1by15T}).
            The SB1 and SB2 templates, however, remain in the same neuron until the size of the map is increased to $1\times20$~(Fig.~\ref{fig: 1by20T}).
            The separation between templates B and E only happens when the size of the SOM exceeds $1\times22$~(Fig.~\ref{fig: 1by22T}).
        \begin{figure*}
            \begin{subfigure}[b]{\textwidth}
                \centering
                \includegraphics[width=\textwidth]{images0.01/1d/dist_1_by_22.png}
            %\caption{$1\times22$ weight map}
             %\label{fig: 1by22T}
            \end{subfigure}
            \hfill
            \begin{subfigure}[b]{\textwidth}
                \includegraphics[width=\textwidth]{images0.01/1d/hit_t_1_by_22.png}
             %\caption{$1\times22$ hits map}
             %\label{fig: 1by22Thits}
            \end{subfigure}
            \caption{The same as Fig.~\ref{fig: 1by2T} but this time the figure shows results of training network in $1\times22$~grid.}
            \label{fig: 1by22T}
        \end{figure*} %PB160427: this figure looks really great! Very easy to understand.

@@

            The fact that the templates need to have at least 22 different neurons to be divided into 12 groups shows that the differences between SB1 and SB2, and B and E, templates are very small.
            Because of their similarities, they tend to stay in the same group until the network becomes big enough to make attention to the smallest particularity.

        \subsubsection{Classifying the galaxy sample}
         \label{sec: 1Dv}
            After training the networks as discussed in Sec.~\ref{sec: 1Dt}, we used them to classify the fitted SEDs of the sample of 142 galaxies from \citetalias{Hossein12}.
            The upper panel in Fig.~\ref{fig: 1by2V} shows the result of this classification using the $1\times2$~network from Fig.~\ref{fig: 1by2T}.
            Eighty of the galaxies have SEDs similar to those of the early type galaxies and the SEDs of the rest are similar to the starburst galaxies.
            %The lower panels show the average SED of the galaxies in each node. 
            The lower left panel shows the average SED of the 80 galaxies that were classified as early type 
            These galaxies are similar to the ones in the left node in the upper panel of Fig.~\ref{fig: 1by2T}:
            the SEDs clearly show the 4000\AA~break, one of the signatures of early type galaxies.
            The H$\alpha$ emission in the spectrum could be from galaxies with similar SEDs to Sa galaxies, but with stronger emission lines.
            The average SED in the lower right panel of Fig.~\ref{fig: 1by2V} shows strong emission lines and bright ultraviolet continuum, indications of a high star formation rate.
            \begin{figure}
                \begin{subfigure}[b]{0.5\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{images0.01/1d/hit_v_1_by_2.png}
                    %\caption{$1\times2$ weight map}
                     %\label{fig: 1by3T}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.5\textwidth}
                     \includegraphics[width=\textwidth]{images0.01/1d/SED_total1by2.png}
                     %\caption{$1\times2$ hits map}
                     %\label{fig: 1by3Thits}
                \end{subfigure}
                \caption{Classification of fitted galaxy SEDs from \citetalias{Hossein12} using the $1\times2$~networks trained from the \citetalias{Kinney96} templates (Fig.~\ref{fig: 1by2T}). Upper panel: a hit map with the number in each node representing the number of galaxies belonging to that group. In this case 80 means that 80 of the 142 SEDs are classified into the early type group while the 62 of the 142 SEDs are categorized as starburst galaxies. Lower panel: Average SED of the galaxies in each group, early-type on the left and starburst on the right.}
                \label{fig: 1by2V}
            \end{figure}

            Since in this network galaxies were forced to be divided into a maximum of two groups, the strongest feature in a galaxy's SED predominantly decides which group the galaxy belongs to.
            Galaxies with a weak 4000\AA~break but strong emission lines and ultraviolet upturn are categorized as starbursts while galaxies with a strong 4000\AA~break are categorized as early type.
            Increasing the size of the SOM helps solve the problem of the galaxies which have features that are common in both groups.

            Fig.~\ref{fig: 1by3V} presents the result of classifying the SED of the galaxies using the $1\times3$~network (from Fig.~\ref{fig: 1by2T}): 66 of the galaxies belong to the early type group, and 47 belong to the starburst group. 
            However, 29 of the galaxies are similar to starbursts in some, but not all, of their features. 
            Galaxies in this group have strong emission lines and are ultraviolet-bright, but they also have a strong 4000\AA~break, which makes them cluster closer to the early type galaxies (middle panel in the lower part of Fig.~\ref{fig: 1by2V}).

@@

            The first two neurons in the upper plot in Fig.~\ref{fig: 1by22V} are empty, while these neurons were occupied by galaxies B and E in the trained network.
            We therefore conclude that there are no galaxies in the \citetalias{Hossein12} sample with SEDs similar to that of type B or E.
            There are 3 galaxies in the third neuron, similar to the S0 type. 
            18 of the galaxies are jn the fifth neurons and similar to the Sa type, while the 7 galaxies in the fourth neurons have SEDs similar to both S0 and Sa galaxies.
            There is one Sb type galaxy and 15 galaxies with SEDs similar to both Sa and Sb type galaxies.
            The following two neurons (eighth and ninth neurons from the left) are on the edge of the early type and starburst galaxies.
            In the upper panel in Fig.~\ref{fig: 1by22T}, the colour between these two neurons is black, which shows that these two are very different.
            Therefore, 24 galaxies in the eighth neurons have similar SEDs to Sb type galaxies, and the SEDs of the 14 galaxies in the ninth neuron are similar to the SB6 galaxies.
            The far right neurons in the network belong to galaxies with SED comparable to types SB1 and SB2.

@@

            %PB20160604: is this paragraph really necessary?
            %%Sr: I think it clears the point of having regions in middle neurons
            %%%%%%%Sr220616: I almost completely changed the following paragraph, but I think it makes more sense, now.
            Other methods for finding the type of galaxies (e.g. $\chi^2$ fitting and supervised neural network) are limited to templates and the outliers are not classified.
            However, using SOM method we categorized galaxies based on their position on trained SOMs, which can have empty neurons.
            The empty neurons in a trained SOM show that there must be another type of template there, that is absent from the K96 template spectra.
            Using the distance of the trained network is the best way to describe the properties of this absent template.
            Therefore, with SOM, we have the freedom to categorize galaxies in intermediate (new) templates.
            %%%%%%%Sr220616: in response to Sophia's comment; In SOM we do not fit the data. we create a network using our template. In the final version of network each neuron has a weight vector. When we give a new data to this weight vector, based on distance of the the weight vector and a new input data, the SOM finds a winner neuron. If this winner neurons were the same for K96 and T12 samples, we would say the input vector in these two case were similar. Since we know type of the galaxies in K96, we could decide about the type of the galaxies in T12.

@@

        In order to check whether our clustering in Sec.~\ref{sec: 1Dv}~are correct, in this section we test the relations between median properties of the galaxies in each neuron.  %%%%%%%Sr220616:changed "conclusion" to clustering!

        \begin{figure}
            \centering
            \includegraphics[width=0.5\textwidth]{images0.01/1d/props5.png}
            \caption{Comparing the median of four properties of galaxies in each node in $1\times2$ (black circles), $1\times4$ (blue diamonds), and $1\times8$ (red triangles) networks.}
            \label{fig: props}
        \end{figure}

        Fig.~\ref{fig: props} shows the median of stellar mass, age, specific star formation rate (sSFR; star formation rate per stellar mass), and $f_\mathrm{burst}$ of the galaxies in each neuron in the $1\times2$, $1\times4$, and $1\times8$ networks.
        In all plots, the horizontal axis is the number of the neurons divided by the size of the network.%, and vertical axis is one of the aforementioned properties.
%        
        As shown in Fig.~\ref{fig: props}, in all three networks, stellar mass and age decrease while sSFR and $f_\mathrm{burst}$ increase as the type changes from early type galaxies to starburst. 
       % These trends are another evidence that this method is working and can be used to categorize galaxies based on their SEDs or even their physical properties. 
        %Considering the results from Fig.~\ref{fig: props}, we can clearly see that s
        Separating galaxies based on SED types also leads to a separation in properties that derived (via {\sc CIGALE}) from the SEDs, as expected since the SED types are also based on {\sc CIGALE} fitting.

@@

               \begin{figure*}
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f1.png}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f2.png}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f3.png}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f4.png}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f5.png}
        \end{subfigure}
       \hfill
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f6.png}
        \end{subfigure}
        \caption{From top to bottom: Correlation between age (t$_{\rm {D4000}}$) and sSFR ($\phi$), sSFR and stellar mass, age (t$_{\rm {D4000}}$) and stellar mass, and A$_{\rm {FUV}}$ with sSFR, age (t$_{\rm {D4000}}$) and stellar mass. Points show the median values of these properties of galaxies in $1\times22$ network and error bars are the standard deviation of the median values. colours shows the type of the galaxies in the new classification. The most pure red colour shows E type galaxies and the most blue one indicates the SB1 galaxies. The other colours depend on how much they are close to red or blue show that the galaxies are early type or starburst.}
        \label{fig: props_vs_props}
    \end{figure*}

        In order to compare our classifications with previous works, we produced similar plots to the ones in both \citetalias{Noll09} and \citetalias{Hossein12} (Fig.~\ref{fig: props_vs_props}).
        In Fig.~\ref{fig: props_vs_props}, the data points are median values of the properties of the galaxies in each neuron in Fig.~\ref{fig: 1by22V}, and error bars show the standard deviation of those properties.
        The colours show the galaxy types, where B type galaxies are shown with red and SB1 types are shown with blue.
        The redder the colour, the more similarity with early type galaxies, while the purple to blue colours show more similarity to starburst galaxies.
        Since in this method we allow galaxies to occupy classes in between the original \citetalias{Kinney96} templates, we have more than 12 colours in Fig.~\ref{fig: props_vs_props}.
        The purple point in the middle of the blue points in  Fig.~\ref{fig: props_vs_props} corresponds to galaxies with SEDs similar to type Sc: the shape of their SED indicates that they are old galaxies with high sSFR.

@@

    \subsection{Two-dimensional self-organizing maps}
    \label{sec: 2D}
    The one-dimensional networks are great tools to categorize the SEDs and monitor the changes of properties of galaxies with category.
    However, each neuron in these networks is limited to be connected to one other neuron in each direction.
    In a 2D network, each neuron can have more than two immediate neighbours, providing a more complete pictures of relations between each SED type.
    As described in Sec.~\ref{sec: method}, one of the main advantages of the SOM is that when the weight of one neuron is adjusted after finding the best matching unit, the weight of the whole map will be changed.
    This quality of the SOMs provides a unique opportunity to analyse 2D networks in two approaches. 
    At first, as in the 1D SOMs, we assumed that all the galaxies must have SEDs similar to any of the 12 templates in Fig.~\ref{fig: k96}.
    In this case, we can categorize SEDs of other galaxies based on those 12 types.
   In the second approach, we trained networks with the assumption that there might be completely different types of galaxies not encountered before.
    This approach can be used to trace and recognize the outliers.
    To execute the second approach, we used the {\tiny selforgmap} library in {\tiny MATLAB}.
    Using the {\tiny selforgmap} code, based on the size of a map and the ordering step neighbourhood distance, we can arrange that the data spread all over the SOM or sit close together, allowing room for outliers.
%    The latter case, can be used to identify outliers.

    To get the best result from both methods, we should have SOMs with ``enough"  neurons to give any new types of galaxies a chance to find their places in the map.
    ``Enough  neurons" is a vague statement, but as mentioned in the Sec.~\ref{sec: method} the size of SOMs is arbitrary and depends on the size of the input data and the kind of information needed from the SOMs.
    Since the training dataset (the \citetalias{Kinney96} templates) has 12 members, we assumed a SOM with size of $8\times8$~would be a good start.
    We then increased the size of the SOM to find the highest size suitable for the training sample.
    We noticed that very few changes occurred after the SOMs were grown to a size greater than $12\times12$, so we used this as the maximum size.

>>>>

    In order to find a sufficient size for the trained networks, we created maps with sizes ranging from $1\times2$ to $50\times50$.
    Varying the grid size of the maps helped us to monitor whether tighter grouping of galaxies happens due to their similar properties or a lack of map space to separate them.
    Based on the size of the data and SOM results, we found the optimum grid size to be $1\times22$ and $12\times12$ in 1D and 2D maps, respectively. 
    For each grid we created different SOMs with different learning factors, neighbourhood distances, and number of iterations to find the most suitable values for our dataset, which are listed in Section~\ref{sec: create_som}. %%%Sr220616: changed the optimum to the most suitable values ... . moved the initial values to section 3.2

@@

    \subsection{One-dimensional self-organizing maps}
    \label{sec: 1D}
        \subsubsection{TRAINING THE NETWORKS}
        \label{sec: 1Dt}
            To start our clustering, we assumed that galaxies can be divided into only two general types: quiescent and starburst.
            This corresponds to a network with only two neurons.
            We increased the size of the map gradually until the 12 input samples divide into the 12 neurons.

            Figs.~\ref{fig: 1by2T} --\ref{fig: 1by22T}~show the results of the training $1\times2$, $1\times3$, and $1\times22$ networks, respectively (the remaining networks are shown in Appendix~\ref{app: high_Z_1d_soms}).
            As in Fig.~\ref{fig: sample}, the upper part of the figures shows the neurons and their relative distances, $D_j$.
            As mentioned in Sec.~\ref{sec: method}, an increase in the darkness of colours between neurons represents an increase in relative distance between the neurons.
            The lower panels of the figures show the number of \citetalias{Kinney96} templates that are placed in each neuron. 
            \begin{figure}
                \begin{subfigure}[b]{0.5\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{images0.01/1d/dist_1_by_2.png}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.5\textwidth}
                     \includegraphics[width=\textwidth]{images0.01/1d/hit_t_1_by_2.png}
                \end{subfigure}
                \caption{Results of training network in $1\times2$~grid. As in Fig.~\ref{fig: sample}, the upper panel is a distance map and the lower panel is a hit map. In this network, 5 of the templates from \citetalias{Kinney96} are categorized as quiescent galaxies and the rest are starbursts. Because of their strong emission lines, Sc galaxies are moved towards the starburst ones.}
                 \label{fig: 1by2T}
            \end{figure}

            In the upper map in Fig.~\ref{fig: 1by2T}, the dark colour between two neurons indicates that the relative distance between these two groups are relatively high, and these two groups are distinguishable groups.
            In the lower part of Fig.~\ref{fig: 1by2T}, we see that the templates are divided into two groups of 5 and 7.
            Although we know from \citetalias{Kinney96} and \citetalias{Hossein12} that 6 of the templates are quiescent and the other 6 are starbursts, the SOM results show 5 of the galaxies in one group and the other 7 in the second group.
            In this method, the Sc template has been categorized as starburst due to the relatively higher disk stellar population compared to that of the bulge. 
            According to \citetalias{Kinney96}, Sc galaxies are considered to be late Hubble type galaxies which have a flatter SED compare to other quiescent galaxies.

            Fig.~\ref{fig: 1by3T} shows the results of the training in a 1$\times$3 network.
            In these plots we force the galaxies to be categorized in a maximum of 3 groups. 
            If the templates in Fig.~\ref{fig: 1by2T} truly belonged in two groups, they would be grouped into two groups in this network, even when we try to cluster them into three. 
            However, in the lower panel of Fig.~\ref{fig: 1by3T}, we can see that the middle node contains two templates.
            These two templates, which are separated from the group of starburst templates in the lower part of Fig.~\ref{fig: 1by2T},  are the SB5 and SB6 types.
            In the upper plot of Fig.~\ref{fig: 1by3T}, the colour between two right neurons is black and the colour between two left neurons is white. 
            The black colour indicates that the left neuron is completely different from the other two groups,
            while the white colour shows that the two right neurons are very similar to one another.

@@

            We increased the size of the maps gradually until the galaxies are divided into twelve groups (Figs.~\ref{fig: 1by4T} to ~\ref{fig: 1by20T} and ~\ref{fig: 1by22T}).
            Since there are more nodes in larger SOMs, the  algorithm 
            pays more attention to small differences between groups.
            If the templates from \citetalias{Kinney96} had completely distinct SEDs, then a $1\times12$-sized SOM would be expected to show 12 different groups each containing a single template.
            However, in Fig.~\ref{fig: 1by12T}, we see that three of the neurons contain two templates and three of them are empty.
            It is evident that there was no template in the \citetalias{Kinney96} sample that can fill those empty neurons.
            In Fig.~\ref{fig: 1by12T}, from left to right, templates with types B and E, SB3 and SB4, and SB1 and SB2 are the ones grouped together. 
            The SB3 and SB4 grouping breaks when we increase the size of the network to $1\times15$~(Fig.~\ref{fig: 1by15T}).
            The SB1 and SB2 templates, however, remain in the same neuron until the size of the map is increased to $1\times20$~(Fig.~\ref{fig: 1by20T}).
            The separation between templates B and E only happens when the size of the SOM exceeds $1\times22$~(Fig.~\ref{fig: 1by22T}).
        \begin{figure*}
            \begin{subfigure}[b]{\textwidth}
                \centering
                \includegraphics[width=\textwidth]{images0.01/1d/dist_1_by_22.png}
            %\caption{$1\times22$ weight map}
             %\label{fig: 1by22T}
            \end{subfigure}
            \hfill
            \begin{subfigure}[b]{\textwidth}
                \includegraphics[width=\textwidth]{images0.01/1d/hit_t_1_by_22.png}
             %\caption{$1\times22$ hits map}
             %\label{fig: 1by22Thits}
            \end{subfigure}
            \caption{The same as Fig.~\ref{fig: 1by2T} but this time the figure shows results of training network in $1\times22$~grid.}
            \label{fig: 1by22T}
        \end{figure*} %PB160427: this figure looks really great! Very easy to understand.

@@

            The fact that the templates need to have at least 22 different neurons to be divided into 12 groups shows that the differences between SB1 and SB2, and B and E, templates are very small.
            Because of their similarities, they tend to stay in the same group until the network becomes big enough to make use of the smallest particularity.

        \subsubsection{Classifying the galaxy sample}
         \label{sec: 1Dv}
            After training the networks as discussed in Sec.~\ref{sec: 1Dt}, we used them to classify the fitted SEDs of the sample of 142 galaxies from \citetalias{Hossein12}.
            The upper panel in Fig.~\ref{fig: 1by2V} shows the result of this classification using the $1\times2$~network from Fig.~\ref{fig: 1by2T}.
            Eighty of the galaxies have SEDs similar to those of the early type galaxies and the SEDs of the rest are similar to the starburst galaxies.
            %The lower panels show the average SED of the galaxies in each node. 
            The lower left panel shows the average SED of the 80 galaxies that were classified as quiescent. 
            These galaxies are similar to the ones in the left node in the upper panel of Fig.~\ref{fig: 1by2T}:
            the SEDs clearly show the 4000\AA~break, one of the signatures of quiescent galaxies.
            The H$\alpha$ emission in the spectrum could be from galaxies with similar SEDs to Sa galaxies, but with stronger emission lines.
            The average SED in the lower right panel of Fig.~\ref{fig: 1by2V} shows strong emission lines and bright ultraviolet continuum, indications of a high star formation rate.
            \begin{figure}
                \begin{subfigure}[b]{0.5\textwidth}
                    \centering
                    \includegraphics[width=\textwidth]{images0.01/1d/hit_v_1_by_2.png}
                    %\caption{$1\times2$ weight map}
                     %\label{fig: 1by3T}
                \end{subfigure}
                \hfill
                \begin{subfigure}[b]{0.5\textwidth}
                     \includegraphics[width=\textwidth]{images0.01/1d/SED_total1by2.png}
                     %\caption{$1\times2$ hits map}
                     %\label{fig: 1by3Thits}
                \end{subfigure}
                \caption{Classification of fitted galaxy SEDs from \citetalias{Hossein12} using the $1\times2$~network trained from the \citetalias{Kinney96} templates (Fig.~\ref{fig: 1by2T}). Upper panel: a hit map with the number in each node representing the number of galaxies belonging to that group. In this case 80 means that 80 of the 142 SEDs are classified into the early type group while the 62 of the 142 SEDs are categorized as starburst galaxies. Lower panel: Average SED of the galaxies in each group, quiescent on the left and starburst on the right.}
                \label{fig: 1by2V}
            \end{figure}

            Since in this network galaxies were forced to divide into a maximum of two groups, the strongest feature in a galaxy's SED predominantly decides which group the galaxy belongs to.
            Galaxies with a weak 4000\AA~break but strong emission lines and ultraviolet upturn are categorized as starbursts while galaxies with a strong 4000\AA~break are categorized as quiescent.
            Increasing the size of the SOM helps solve the problem of the galaxies which have features common to both groups.

            Fig.~\ref{fig: 1by3V} presents the result of classifying the SED of the galaxies using the $1\times3$~network (from Fig.~\ref{fig: 1by2T}): 66 of the galaxies belong to the early type group, and 47 belong to the starburst group. 
            However, 29 of the galaxies are similar to starbursts in some, but not all, of their features. 
            Galaxies in this group have strong emission lines and are ultraviolet-bright, but they also have a strong 4000\AA~break, which makes them cluster closer to the quiescent galaxies (middle panel in the lower part of Fig.~\ref{fig: 1by2V}).

@@

            The first two neurons in the upper plot in Fig.~\ref{fig: 1by22V} are empty, while these neurons were occupied by galaxies B and E in the trained network.
            We therefore conclude that there are no galaxies in the \citetalias{Hossein12} sample with SEDs similar to that of type B or E.
            There are 3 galaxies in the third neuron, similar to the S0 type. 
            Eighteen of the galaxies are in the fifth neuron and similar to the Sa type, while the 7 galaxies in the fourth neuron have SEDs similar to both S0 and Sa galaxies.
            There is one Sb type galaxy and 15 galaxies with SEDs similar to both Sa and Sb type galaxies.
            The following two neurons (eighth and ninth neurons from the left) are on the edge of the early type and starburst galaxies.
            In the upper panel in Fig.~\ref{fig: 1by22T}, the colour between these two neurons is black, which shows that these two are very different.
            Therefore, 24 galaxies in the eighth neurons have similar SEDs to Sb type galaxies, and the SEDs of the 14 galaxies in the ninth neuron are similar to the SB6 galaxies.
            The far right neurons in the network belong to galaxies with SED comparable to types SB1 and SB2.

@@

            %PB20160604: is this paragraph really necessary?
            %%Sr: I think it clears the point of having regions in middle neurons
            %%%%%%%Sr220616: I almost completely changed the following paragraph, but I think it makes more sense, now.
            Other methods for finding the type of galaxies (e.g. $\chi^2$ fitting and supervised neural network) are limited to templates: the outliers cannot be classified.
            However, using the SOM method we categorized galaxies based on their position on trained maps, which can have empty neurons.
            The empty neurons in a trained SOM show that there must be another type of template that is absent from the \citealias{K96} template spectra.
            Using the distance within the trained network is the best way to describe the properties of this absent template.
            Therefore, with SOM, we have the freedom to categorize galaxies in intermediate (new) templates.
            %%%%%%%Sr220616: in response to Sophia's comment; In SOM we do not fit the data. we create a network using our template. In the final version of network each neuron has a weight vector. When we give a new data to this weight vector, based on distance of the the weight vector and a new input data, the SOM finds a winner neuron. If this winner neurons were the same for K96 and T12 samples, we would say the input vector in these two case were similar. Since we know type of the galaxies in K96, we could decide about the type of the galaxies in T12.

@@

        In order to check whether our clustering in Sec.~\ref{sec: 1Dv}~is correct, in this section we examine the relations between median properties of the galaxies in each neuron.  %%%%%%%Sr220616:changed "conclusion" to clustering!

        \begin{figure}
            \centering
            \includegraphics[width=0.5\textwidth]{images0.01/1d/props5.png}
            \caption{Median values of four properties of galaxies in each node in $1\times2$ (black circles), $1\times4$ (blue diamonds), and $1\times8$ (red triangles) networks.}
            \label{fig: props}
        \end{figure}

        Fig.~\ref{fig: props} shows the median values of stellar mass, age, specific star formation rate (sSFR; star formation rate per stellar mass), and $f_\mathrm{burst}$ of the galaxies in each neuron in the $1\times2$, $1\times4$, and $1\times8$ networks.
        In all plots, the horizontal axis is the number of the neurons divided by the size of the network.%, and vertical axis is one of the aforementioned properties.
%        
        As shown in Fig.~\ref{fig: props}, in all three networks, stellar mass and age decrease while sSFR and $f_\mathrm{burst}$ increase as the type changes from early type galaxies to starburst. 
       % These trends are another evidence that this method is working and can be used to categorize galaxies based on their SEDs or even their physical properties. 
        %Considering the results from Fig.~\ref{fig: props}, we can clearly see that s
        Separating galaxies based on SED types also leads to a separation in properties that derived (via {\sc CIGALE}) from the SEDs, as expected since the SED types are also based on {\sc CIGALE} fitting.

@@

               \begin{figure*}
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f1.png}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f2.png}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f3.png}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f4.png}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f5.png}
        \end{subfigure}
       \hfill
        \begin{subfigure}[b]{0.3\textwidth}
            \includegraphics[width=\textwidth]{images0.01/1d/f6.png}
        \end{subfigure}
        \caption{From top to bottom: Correlation between age (t$_{\rm {D4000}}$) and sSFR ($\phi$), sSFR and stellar mass, age (t$_{\rm {D4000}}$) and stellar mass, and A$_{\rm {FUV}}$ with sSFR, age (t$_{\rm {D4000}}$) and stellar mass. Points show the median values of these properties of galaxies in the $1\times22$ network and error bars are the standard deviation of the median values. Colours show the type of galaxies in the new classification. The purest red colour shows E type galaxies and the bluest one indicates the SB1 galaxies. The other colours are determined by the galaxy group's relative distance to these two endpoints.}
        \label{fig: props_vs_props}
    \end{figure*}

        In order to compare our classifications with previous works, we produced similar plots to the ones in both \citetalias{Noll09} and \citetalias{Hossein12} (Fig.~\ref{fig: props_vs_props}).
        In Fig.~\ref{fig: props_vs_props}, the data points are median values of the properties of the galaxies in each neuron in Fig.~\ref{fig: 1by22V}, and error bars show the standard deviation of those properties.
        The colours show the galaxy types, where B type galaxies are shown with red and SB1 types are shown with blue.
        The redder the colour, the more similarity with  quiescent galaxies, while the purple to blue colours show more similarity to starburst galaxies.
        Since in this method we allow galaxies to occupy classes in between the original \citetalias{Kinney96} templates, we have more than 12 colours in Fig.~\ref{fig: props_vs_props}.
        The purple point in the middle of the blue points in  Fig.~\ref{fig: props_vs_props} corresponds to galaxies with SEDs similar to type Sc: the shape of their SED indicates that they are old galaxies with high sSFR.

@@

    \subsection{Two-dimensional self-organizing maps}
    \label{sec: 2D}
    One-dimensional networks are great tools to categorize the SEDs and monitor the changes of properties of galaxies with category.
    However, each neuron in these networks is limited to be connected to one other neuron in each direction.
    In a 2D network, each neuron can have more than two immediate neighbours, providing a more complete pictures of relations between SED types.
    As described in Sec.~\ref{sec: method}, one of the main advantages of the SOM is that when the weight of one neuron is adjusted after finding the best matching unit, the weight of the whole map will be changed.
    This quality of the SOMs provides a unique opportunity to analyse 2D networks in two approaches. 
    At first, as in the 1D SOMs, we assumed that all the galaxies must have SEDs similar to any of the 12 templates in Fig.~\ref{fig: k96}.
    In this case, we can categorize SEDs of other galaxies based on those 12 types.
   In the second approach, we trained networks with the assumption that there might be completely different types of galaxies not encountered before.
    This approach can be used to trace and recognize the outliers.
    To execute the second approach, we used the {\tiny selforgmap} library in {\tiny MATLAB}.
    Using the {\tiny selforgmap} code, based on the size of a map and the ordering step neighbourhood distance, we can arrange that the data spread all over the SOM or sit close together, allowing room for outliers.
%    The latter case, can be used to identify outliers.

    To get the best result from both methods, we should have SOMs with ``enough"  neurons to give any new types of galaxies a chance to find their places in the map.
    ``Enough  neurons" is a vague statement, but as mentioned in Sec.~\ref{sec: method} the size of SOMs is arbitrary and depends on the size of the input data and the kind of information needed from the SOMs.
    Since the training dataset (the \citetalias{Kinney96} templates) has 12 members, we assumed a SOM with size of $8\times8$~would be a good start.
    We then increased the size of the SOM to find the highest size suitable for the training sample.
    We noticed that very few changes occurred after the SOMs were grown to a size greater than $12\times12$, so we used this as the maximum size.
